The Spatial and Temporal Domains of Modern Ecology

### 2017-06-20 paper/main.tex
Full revision, including discussion and NatEcolEvol style abstract.

%\begin{sciabstract}
\section*{Abstract}
%Ecologists have long known that the observational scales play a critical role in shaping our understanding of ecosystems, and that observations spanning multiple scales of space and time are needed to effectively characterize ecological phenomena. 
\textbf{In order to properly understand ecological phenomena, it is necessary to quantify their behavior over the range of spatial and temporal scales at which they manifest. Ecology has been concerned with this need for decades, and the ability to collect ecological observations across multiple scales has grown rapidly since then. Characterizing the spatial and temporal domains (the distribution of observations within one or more scale spectra) of modern ecological observations can therefore provide important insight into the field's progress towards a more comprehensive understanding of ecosystem behaviour. To characterize these domains, we conducted a meta-analysis of recent (2004-2014) ecological studies, in which we quantified four primary dimensions of their reported observations: plot resolution, sampling interval, effective duration (time between start and end of temporal replicates), and effective extent (area enclosed by spatial replicates). We also estimated the \emph{actual} extent and duration, which respectively represent the summed area and time covered by spatial and temporal replicates. \textcolor{blue}{Replace this text with more specific summary of observed scales:} \textcolor{red}{Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}}}. % Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations). Resolution, extent, and duration were not clearly reported in most reviewed studies. This indicates inattentiveness to scale, which, together with the narrow spatio-temporal domains of observations, suggests a limited ability to generalize the knowledge gained in many ecological studies, which is critical for developing an ability to predict ecological responses in an era of increasing global change \cite{levin_problem_1992}. An emerging wave of observing and analytical technologies (e.g. nano-satellites, unmanned systems, open image archives, and high-power processing platforms) now makes it dramatically easier and cheaper to collect high-quality, ecologically meaningful measurements at scales inaccessible to field scientists. Combined with a greater attention to scale in research design, these advances can facilitate the spread of ecological research to sparsely observed scales, while improving the transferability of ecological knowledge.}

@@

\noindent \textbf{\emph{Domains}}\\
In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1-10,000 ha, 19\% 10,000-100,000 ha, 27\% 100,000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig. 1B).

@@

Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of ecological observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (the 37\% that were unreplicated were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

To provide further insight into observational domains, we also evaluated the degree to which the scales \emph{represented} by observations differed from those that were actually observed. To make this assessment, we first log$_10$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, in order to calculate the magnitude of difference (or decade) between each pair of dimensions for each observation. We then compared how these magnitudes varied in relation to the scale of the actual dimension (Fig. 3). This comparison showed that a majority (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).  Actual extent converged with extent $\geq$1 100,000 ha, but this applies to just 5\% of observations. The actual duration of 64\% of observations was $<$ 1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig. 3B). The two duration measures were approximately equal for the 16\% of observations exceeding one month of actual duration.

@@

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates. Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation). Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This somewhat clearer trend also suggests that including more recent studies would show extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).

@@

%This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI). Observations that were non-continuous in time or space (e.g. point-based field measurements) may effectively represent larger scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}. From a spatial perspective, this concern does not apply to contiguous sampling schemes, which were primarily based on remote sensing and cover the largest area, thus the net effect is that this analysis may underestimate the effective areas of smaller observational extents. From a temporal perspective, our analysis likely underestimates the effective duration of many observations, particularly those where instantaneous, repeated measures of slow-changing ecological features were made. Snapshots in time may be sufficient to capture the temporal dynamics of such phenomena (e.g. changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these the total time elapsed between the first and last measurements (the span of observation) may more closely approximate effective duration. However, many studies focused on more dynamic phenomena, and for these long periods of continuous observation may be more important for understanding dynamics than frequent repeats. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}, except patterns resolving themselves at scales greater than 12-24 hours. For this reason, and to provide a consistent standard for estimating duration, we recorded actual duration.

%Adding to our motivation to record actual observational scales was the fact that

\noindent \textbf{Broader implications}\\
Our results suggest that the spatial and temporal domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques. Spatially, the majority of observations have grains of $<1$ m$^2$ and extents of $<$10,000 ha (Fig 1A;B). In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig. 1C) and relatively short duration ($leq$1 year; Fig. 1D; 2D). Comparisons between observational dimensions revealed that larger extents and are associated with larger plot sizes (Fig. 2C), while longer durations are associated with longer intervals (Fig. 2D). The latter association presumably reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of field-based data collection methods. This same tradeoff is also responsible for the inverse relationship between resolution and interval that is the most obvious feature within that domain space (Fig. 1A).

Comparing these general findings to those of earlier related studies, it appears that modern ecology's observational scales have not changed greatly in the past few decades.

These general results suggest that the scale domains of modern ecology have not changed greatly since the 1980s.  Porter's study.

Comparing these primary tendencies to earlier efforts, it seems that not much has changed in ecological

Taken collectively, these

\begin{enumerate}
\item General patterns of scales 
  \subitem{-} General patterns
  \subitem{-} The continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
  \subitem{-} Notable holes: high frequency (sub-daily, daily), high resolution 
   \subitem{-} Change since Tilman/Kareiva
   \subitem{-} Duration hasn't improved much since Tilman (check this).
   \subitem{-} Evaluate how duration has improved since then. 387/749 of the papers he looked at from the journal Ecology (1977-1984) were mainly experimental or had an experimental component. 99 papers were mainly theoretical or conceptual. Remainder observational. We could compare our results to that here. Of his 623 field based studies (including 180 experiments), $\sim$41\% lasted less than one year,  $\sim$85\% less than three years, most studies $>$10 years and all $>$50 years were paleoecology/chronosequences.
  \subitem{-} Maybe mention here (or probably better later in the discussion), that our decision has the unfortunate consequence of not being directly comparable to previous assessments that included experiments. More of just a caveat or note for the reader who is familiar with the earlier work and their summary statistics
  \subitem But there do seem to be trends increasing extent (and duration, but need to put this above), perhaps due to increasing use of remote sensing 
\item Scales are not well-documented 
  \subitem Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}. (Scale is not a concern that cuts across the entire discipline of ecology).

\item Represented scale is very different from actual scale of observations. 
  \subitem In heeding Levin's call to think at multiple and appropriate ecological scales, we also need to give pause to think about whether our data are overextrapolated/overleveraged.
  \subitem How over-leveraged/under-leveraged might depend on the auto-correlation properties of feature being observed. Maybe use example of fire process
    \subitem Maybe we should go further here to explain why this is an important (and overlooked) assessment to make (actual versus effective comparison)
\item Where to from here?  What do we need to improve?   
  \subitem Satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.

\end{enumerate}

@@

%These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.

Narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) 
Ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}, 
a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.

Our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research. Our results show that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month). Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989}. Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}. Taken together, this narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that the ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}, a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.

The generally small spatial scales of observation is a consequence of the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing. Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this method has not yet been widely adopted in ecological research. Two reasons lie behind this slow uptake. First, remote sensing can be a challenging method for ecologists to learn, many of whom may not have access to appropriate training \cite{pettorelli_satellite_2014}. Second, satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.

In contrast to remote sensing, ecologists have made broader use of technologies that increase the temporal resolution of observations. Automated sensors were used to record 12\% of all observations, and accounted for most very high frequency measurements (intervals $\leq$1 hour). As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas longer-interval data (e.g. annual biomass accumulation) cannot be disaggregated into shorter interval measurements (e.g. weekly biomass accumulation) without interpolation, which turns data into modeled, rather than direct, observations.

In the coming years, rapid technological advances should increase the concentration of ecological observations in currently under-represented domains. The growing numbers of high-resolution satellite sensors, together with new analytical platforms that provide free access to large volumes of pre-processed data and computational power \cite{googleearthengine}, will lower technical barriers that have so far prevented ecologists from adopting this observational technology  \cite{pettorelli_satellite_2014}. Similarly, the advent of unmanned systems offers the ability to measure ecological features at high spatial and temporal frequencies over large areas \cite{anderson_lightweight_2013}, which were scales that were previously impractical to access. The ever-falling cost of sensor technology and the ubiquity of cell phones also means that ecologists, together with a growing army of citizen-scientists, have the unprecedented ability to make spatially dense, high frequency observations over large areas \cite{wolf_gsm-based_2012,collins_new_2006,porter_wireless_2005,dickinson_current_2012}.  A greater attentiveness to scale in general, including more meticulous documentation of observed dimensions, may help to facilitate the spread of ecological research to sparsely studied scales, while improving transferability of knowledge within the discipline.

%Plot resolution issue: need high resolution over large extents, because this relies on remote sensing, and to really capture details, need high resolution.  
%likely because of perceived (and in some cases real) logistical barriers \cite{pettorelli_satellite_2014}.

%These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989}.

%Despite the understanding that ecological patterns and process change with scale \cite{levin_problem_1992}, the narrow distribution of observations within spatial domains calls into question the degree to which ecological phenomena are being investigated at multiple scales. Adding to this concern is the fact the scales of observation were not clearly reported in the majority of studies; the spatial resolution and extent of studies was uncertain in 64\% and 70\% of cases, respectively, while the precise interval and study duration were uncertain in 43\% and 86\% of cases. Although sensitivity analyses show that our conclusions are robust to this uncertainty (SI), the unclear documenting of observational scales suggest that they are not a primary methodological concern in ecological research \cite{chave_problem_2013, wheatley_factors_2009}. 
%
%On the other hand, we found a greater diversity of scales in the temporal domains of observations (Fig. 1D). Over 70\% of sampled observations involved repeated assessments, and one-quarter of these were made with high frequency (intervals $\leq$1 day). Automated, surface-based sensors were used to record most of the very high frequency observations (intervals $\leq$1 hour), and comprised 9\% of all observations, suggesting that ecologists are adopting new technologies for collecting temporal data more rapidly than they are for spatial data.  As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses, whereas it is extremely difficult, if not impossible, to disaggregate long-interval data. 
%
%The temporal duration of most collected observations appears to be brief ($\leq$1year), which is due to the fact that we estimated the summed time of repeated observations, rather than the total time elapsed between the first and last measurements (the span of observation). In many cases, the actual observations are very brief, and thus the temporal duration is narrow. Snapshots in time may be sufficient to capture the temporal dynamics of many phenomena (e.g. changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these our measurement of this domain may understate the temporal breadth of ecological observations. However, for more dynamic phenomena, long periods of continuous observation may be more important for understanding dynamics than frequent repeats. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe behavior and evolution of the active fire \cite{clements_observing_2007}. For this reason, and because duration is the temporal analog of spatial extent, we chose to record observational duration rather then span.

%Based on the observational domains recorded here, it appears that ecologists are still primarily collecting data at very fine spatial scales and over very narrow extents. Although then temporal duration of most ecological observations is also narrow, and a sizable number of observations are not repeated, greater effort is being devoted to capturing temporal dynamics. These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989}.

%Based on this analysis, ecology thus appears to have

%High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
%Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).

%Although limited sample size, insight into the scales at which science of ecology is making bulk of observations

>>>>

%\begin{sciabstract}
\section*{Abstract}
%Ecologists have long known that the observational scales play a critical role in shaping our understanding of ecosystems, and that observations spanning multiple scales of space and time are needed to effectively characterize ecological phenomena. 
\textbf{To properly understand ecological phenomena, it is necessary to observe them across a range of spatial and temporal scales. Ecologists first raised this point in the 1980s, since when the ability to collect multi-scale observations has grown rapidly. To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 studies published between 2004-2014. We found that the scale domains of observations remain fairly narrow, and are collected primarily with conventional field techniques. Compared to prior meta-analyses from the 1980s and early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger. Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, raising questions about observational representativeness. Adding to this concern, observational scales were not clearly reported in most studies, suggesting that scale often is a minor consideration. Journals can help mitigate this problem by adopting scale reporting standards, which can further spur ecologists to adopt rapidly improving observational technologies, and thereby close key gaps in current observational domains.}

%as well as \emph{actual} extent (the summed resolution of spatial replicates) and \emph{actual} duration (integrated sample duration of temporal replicates). Our results suggest that the scale domains of modern ecological observations remain fairly narrow, and are collected primarily with conventional field techniques. A comparison with results from prior earlier, scale-focused meta-analyses revealed that observational durations and resolutions are largely unchanged since the 1980s, but intervals have become finer and extents larger, the latter driven by steadily increasing use of remote sensing.
%Most observations had resolutions of $\leq$1 m$^2$ (67\%) and extents $\leq$10,000 ha (53\%). Temporally, most observations were either unreplicated (37\%) or had intervals $\geq$1 month (26\%), while 64\% of observations had durations of $\leq$1 year. Juxtaposing dimensions revealed several notable ``holes'' (domains with few to no observations), particularly in the realms of high frequency (sub-daily to daily intervals) observations with fine to moderate resolutions ($\geq$1 m$^2$ up to 100 ha) and $\geq$decadal durations. These holes relate to tradeoffs between observational frequency and resolution, as well as duration, that are characteristic of field-based data collection, which was used for 80\% of observations. An analysis of values reported in the most comparable prior scale meta-analyses revealed that observational durations and resolutions are largely unchanged since the 1980s, but intervals have become finer and extents larger. Our data also showed an internal trend of increasing observational extent, coinciding with a 1.3\% per year increase in remote sensing-based observations. Despite this trend, the \emph{actual} extent of 81\% of observations was 4-8 orders of magnitude smaller than extent. Similarly, the \emph{actual} duration of 67\% of observations was 3-9 orders of magnitude shorter than duration. These large discrepancies raise questions about the representativeness of observations, which may in part be addressed by examining the degree to which auto-correlation processes close the gap between actual and represented scales of observation. Adding to this concern is the fact that most observations did not clearly report observational scales, suggesting that scale remains a minor consideration throughout much of ecology.}

@@

\noindent \textbf{Results}\\
\noindent \textbf{\emph{Domains}}\\
In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1-10,000 ha, 19\% 10,000-100,000 ha, 27\% 100,000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig. 1B).

@@

Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

To provide further insight into observational domains, we also evaluated the degree to which the scales \emph{represented} by observations differed from those that were actually observed. To make this assessment, we first log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, in order to calculate the magnitude of difference (or decade) between each pair of dimensions for each observation. We then compared how these magnitudes varied in relation to the scale of the actual dimension (Fig. 3). This comparison showed that a majority (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).  Actual extent converged with extent $\geq$1 100,000 ha, but this applies to just 5\% of observations. The actual duration of 64\% of observations was $<$ 1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig. 3B). The two duration measures were approximately equal for the 16\% of observations exceeding one month of actual duration.

@@

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates. Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation). Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This somewhat clearer trend also suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).

@@

%This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI). Observations that were non-continuous in time or space (e.g. point-based field measurements) may effectively represent larger scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}.

%Adding to our motivation to record actual observational scales was the fact that 
\noindent \textbf{Discussion}\\
\noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
Our results suggest that the scale domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques. Spatially, the majority of observations have grains of $\leq$1m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B). In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig. 1C) and relatively short duration ($\leq$1 year; Fig. 1D; 2D). Contrasting observational dimensions reveals that larger extents are associated with larger plot sizes (Fig. 2C), while longer durations are associated with longer intervals (Fig. 2D). The latter association presumably reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation. The same tradeoff is also responsible for the inverse relationship between resolution and interval that dominates that domain space (Fig. 2A). As a result of these tradeoffs, there are notable ``holes'' in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig. 2A) and 2) decadal or longer durations (Fig. 2B).

%there are prominent holes in the scale domains of ecology, notably in the spaces defined by observations having 1) daily to sub-daily intervals and resolutions $\geq$1 m$^2$ up to 100 ha; 2) decadal or longer durations and extents $\geq$10,000,000 ha, such that few observations are collected at 1) high spatial and temporal resolution (1 m$^2$, 2) over large areas for long time periods.

Have these domains changed since the seminal papers on scale first began to appear in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}? A comprehensive answer to this question is not possible, absent a similar study targeting earlier literature, but an analysis of results presented in three prior studies provides partial insight. The first and most comprehensive dataset consists of duration values extracted by Tilman \cite{tilman_ecological_1989} from 623 studies published between 1977-1987 in the journal \emph{Ecology}. The average duration of the most comparable subset of those values (n=419; see SI) was 3.6 years, compared to 3.3 years for observations in our sample (or 5.1 if temporally un-replicated observations are excluded). The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who present the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986. The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,510,247 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (197 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments. The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}). The mean interval was 178 days, compared to 684 in our sample, but the 80th percentile value in our study was 169 compared to 329 in the earlier study. Extent in our sample was substantially larger according to multiple summary statistics, including the mean (115,098,283 ha in our study versus 368,403 ha), median (5,120 ha versus 9 ha), and 95th percentile (45,104,594 ha versus 136,000 ha).

Although limited due to methodological differences (e.g. a focus on experiments versus un-manipulated systems; SI), these comparisons suggest that the duration and, less clearly, resolution of ecological observations have changed little in the past 30 years, but observational frequency and particularly extent have increased. The growth in observational extent is also reflected by the positive year-on-year trend within our own dataset, which corresponds to the increasing use of remote sensing (SI).

However, even though observational extent is increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates purportedly represent (Fig. 3A). A similarly large discrepancy is also evident between the time spanned by repeated observations and the time that is spent actually observing a phenomenon (Fig. 3B). These differences between the actual and \emph{representative} scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}. Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (as an example with respect to temporal replication, annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe fire behavior \cite{clements_observing_2007}. To capture such behavior, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.

It is therefore important to examine whether the phenomena being observed are adequately captured by the design of replicates. Our methods suggest one possible procedure for assessing the representativeness of replicates: 1) measure auto-correlation (spatial or serial) in the replicates, 2) add the autocorrelation length to the replicate area/duration, 3) calculate an autocorrelation-adjusted actual extent/duration, and 4) plot where it falls between actual extent/duration and extent/duration. If the adjusted actual value is substantially closer to the representative value, then this may provide greater confidence in the adequacy of the sampling scheme. If not, then alternative sampling methods may be used to close this gap. For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent. Similarly, the interval of high-resolution (a desirable image property because it allows individual ecological features to be more effectively discerned  \cite{dark_modifiable_2007, hay_comparison_2003}) imaging is now approaching daily to sub-daily scales \cite{drusch_sentinel-2:_2012, hand_startup_2015}, allowing improved representation of temporal dynamics. For phenomena that can't be measured from space, either because they are not visible or because they require continuous observation, new approaches for \emph{in situ} or near-surface observations (e.g. low-cost wireless sensors \cite{wolf_gsm-based_2012, collins_new_2006, porter_wireless_2005}, citizen observers \cite{dickinson_current_2012}, and autonomous vehicles \cite{anderson_lightweight_2013}) can be used to increase the spatial and temporal coverage of observations.

The aforementioned insights regarding modern observational domains must be tempered by the uncertainty within our own scale estimates, as detailed in the preceding section. However, most of this uncertainty is attributable to unclear reporting of scale values in the majority of papers we reviewed (a problem also noted in geography studies \cite{margulies_ambiguous_2016}). This tendency towards vague documentation offers one final insight, which is that, despite decades of accumulated knowledge regarding the importance of scale in ecology \cite{levin_problem_1992, wiens_spatial_1989, chave_problem_2013, wheatley_factors_2009}, scale appears to remain a low priority throughout much of the discipline. Beyond contributing to the broader problem of scientific reproducibility \cite{baker_1500_2016}, inattentiveness to scale increases the risk that observations inadequately represent the phenomenon of interest, thereby limiting the generalizability of any derived ecological knowledge \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}. To mitigate this problem, we recommend that ecological journals require authors to quantify and clearly report the values of resolution, extent, interval, and duration.

\noindent \textbf{\emph{Looking forward}}\\
Our study suggests that the concept of scale has yet to fully permeate the discipline of ecology. Evidence for this assertion lies in the continued narrowness of ecology's observational scale domains and the poor documentation of scale dimensions in the literature. However, the increasing extent of ecological observations, enabled by remote sensing and presumably motivated by many ecologists' appreciation of scale-related issues, suggests that ecology's scale domains are gradually changing. In the coming years, the accelerating gains in technology and analytical methods will allow researchers new and unprecedented capabilities to peer into, and thus close, the prominent holes in observational scale domains. A renewed, discipline-wide focus on scale's importance, including the adoption of stricter scale-reporting standards by journals, will help spur ecologists to address these gaps, while fostering the improved transferability of knowledge within the discipline.

@@

%These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.
%
%Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989}. 
%
%a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}. 
%
%As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas longer-interval data (e.g. annual biomass accumulation) cannot be disaggregated into shorter interval measurements (e.g. weekly biomass accumulation) without interpolation, which turns data into modeled, rather than direct, observations.

%In the coming years, rapid technological advances should increase the concentration of ecological observations in currently under-represented domains. The growing numbers of high-resolution satellite sensors, together with new analytical platforms that provide free access to large volumes of pre-processed data and computational power \cite{googleearthengine}, will lower technical barriers that have so far prevented ecologists from adopting this observational technology  \cite{pettorelli_satellite_2014}. Similarly, the advent of unmanned systems offers the ability to measure ecological features at high spatial and temporal frequencies over large areas \cite{anderson_lightweight_2013}, which were scales that were previously impractical to access. The ever-falling cost of sensor technology and the ubiquity of cell phones also means that ecologists, together with a growing army of citizen-scientists, have the unprecedented ability to make spatially dense, high frequency observations over large areas \cite{wolf_gsm-based_2012,collins_new_2006,porter_wireless_2005,dickinson_current_2012}.  A greater attentiveness to scale in general, including more meticulous documentation of observed dimensions, may help to facilitate the spread of ecological research to sparsely studied scales, while improving transferability of knowledge within the discipline.
